---
title: "Tutorial 1 Assignment Code"
output: html_document
---

# BEFORE YOU START 

Create a new RStudio Project for Tutorial 1 and copy this .Rmd file into it.

## Packages

Install the required packages *webchem* and *tidyverse*. We will use *webchem* to quickly access the online chemical database **PubChem** and convert our compound list into a format recognized by CompTox.

```{r packages, message = FALSE, error = TRUE}
# if(!require("webchem")) install.packages("webchem") #Only runs if the package is not already installed on your device.
# if(!require("tidyverse")) install.packages("tidyverse")
# if(!require("ggplot2")) install.packages("ggplot2")

#old version or Jupyterhub
#install_version("Rcpp", version = "1.0.3", repos = "http://cran.us.r-project.org")


list.of.packages <- c("webchem", "tidyverse", "ggplot2", "tidyr", "Rcpp", "ggrepel", "gridExtra",  "gridExtra", "magrittr", "dplyr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(webchem) #Loads package into memory so it can be used by the code.
library(tidyverse)
library(ggplot2)

#library(gridExtra)
#library(Rcpp)
library(tidyr)
library(ggrepel)


library(magrittr) # needs to be run every time you start R and want to use %>%
library(dplyr)    # alternatively, this also loads %>%

# 
# if (!require("ggrepel")) {
#     install.packages("ggrepel", repos="http://cran.fhcrc.org")
#     library(ggrepel)
# }
# 
# if (!require("ggplot2")) {
#     install.packages("ggplot2", repos="http://cran.fhcrc.org")
#     library("ggplot2")
# }
# 
# if (!require("gridExtra")) {
#     install.packages("gridExtra", repos="http://cran.fhcrc.org")
#     library("gridExtra")
# }


```

# Part 1: Data Analysis

Create SMILES for each of the FTOH and PFAA structures provided in the Tutorial 1 - Week 2 slides. I strongly recommend writing them in order of increasing chain length, as this will help you interpret your figures later. You can copy and paste the SMILES strings (without the quotation marks) directly into ChemDraw to confirm whether you made them correctly.

```{r}
cmpdList1 <- c( #Perfluoroalkyl acids
  "FC(F)(F)C(F)(F)C(F)(F)C(=O)O", #PFBA
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(=O)O", #PFPA
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(=O)O", #PFHxA
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(=O)O", #PFHpA
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(=O)O", #PFOA
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(=O)O", #PFNA
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(=O)O", #PFDA
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(=O)O", #PFUA
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(=O)O", #PFDoDA
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(=O)O" #PFTrDA
)

# Every SMILES string must be enclosed in quotation marks.
# Also make sure that every line inside the brackets ends in a comma EXCEPT the last one.


cmpdList2 <- c( #Fluorotelomer alcohols
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)CCO", #4:2 FTOH
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)CO", #5:1 FTOH
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)CCO", #6:2 FTOH
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)CO", #7:1 FTOH
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)CCO", #8:2 FTOH
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)CO", #9:1 FTOH
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)CCO", #10:2 FTOH
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)CO", #11:1 FTOH
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)CCO", #12:2 FTOH
  "FC(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)C(F)(F)CO" #13:1 FTOH
)


```


The code below will convert the SMILES strings into a search format recognized by the CompTox database. After running this block you should have two text files in your project folder.

```{r export}
compounds <- tibble(SMILES = cmpdList1) #Converts the input data into a 1-column tibble (tidy table)

PubChem_IDs <- get_cid(compounds$SMILES, from = "smiles") #Obtains the PubChem database IDs for each input compound

InChiKeys <- pc_prop(PubChem_IDs$cid, properties = c("InChIKey", "MolecularFormula")) %>% #Obtains the International Chemical Identifier Key (InChIKey) for each PubChem ID
  mutate(CID = as.character(CID))

write_lines(InChiKeys$InChIKey, file = "copy_into_CompTox1.txt") #Outputs to a tab-delimited text file. Feel free to rename this file.



compounds2 <- tibble(SMILES = cmpdList2) 

PubChem_IDs2 <- get_cid(compounds2$SMILES, from = "smiles") 

InChiKeys2 <- pc_prop(PubChem_IDs2$cid, properties = c("InChIKey", "MolecularFormula")) %>% 
  mutate(CID = as.character(CID))

write_lines(InChiKeys2$InChIKey, file = "copy_into_CompTox2.txt") #Rename this text file if desired.
```


# Part 2: CompTox

Step 1: Open the CompTox Dashboard in your internet browser: https://comptox.epa.gov/dashboard/batch-search

Step 2: Under Input Types, select **InChIKey**.

Step 3: Open one of the text files containing your converted compound list and copy the entire contents into the Search window. Make sure there are no empty lines.

Step 4: Below the Search window, click **Choose Export Options**.

Step 5: Scroll down, and use the dropdown to select **CSV** as your output format.

Step 6: Select **OPERA Model Predictions** (found in the Intrinsic and Predicted Properties category). Don't click any other options, except those already highlighted.

Step 7: At the bottom of the page, click **Download Export File**. A popup will appear. Wait a few seconds for it to finish processing, then click **Download File**.

Step 8: Copy the downloaded CSV file into your project folder. Either rename your file to "comptoxresults1.csv"/"comptoxresults2.csv", or change the first parameter of *read_csv* (below) to match the name of your file.

Step 9: Repeat Steps 3-8 for the other list of compounds.


When you've completed the above steps, you can go ahead and run the next code block to process the results into something more manageable.

```{r import, message = FALSE}
compTox <- read_csv("comptoxresults1.csv", col_names = TRUE) %>%
  select(matches("INPUT|NAME|AOH|BIO|HENRY|KM|KOC|LOG")) %>% #Uses text matching to keep only the columns we care about
  rename_with(~ c("Key","Name", "AOH", "BCF", "BIODEG_HL", "HenryLaw_Const", "KM", "log_KOA", "KOC", "log_KOW")) %>% #Renames the columns to be a little more readable
  rowwise() %>%
  mutate(log_KAW = (log_KOW - log_KOA), log_KOC = log10(KOC)) %>%
  select(-c(HenryLaw_Const, KOC, AOH, BCF, BIODEG_HL, KM)) %>% #Deletes unwanted columns
  ungroup() %>%
  mutate(Group = "Group1", CmpdID = row_number()) %>% 
  relocate(matches("Group|CmpdID"), .before = Name)

results <- compounds %>%
  full_join(PubChem_IDs, by = c("SMILES" = "query")) %>% #Combines previously created tables by similar columns
  full_join(InChiKeys, by = c("cid" = "CID")) %>%
  full_join(compTox, by = c("InChIKey" = "Key"))





compTox2 <- read_csv("comptoxresults2.csv", col_names = TRUE) %>%
  select(matches("INPUT|NAME|AOH|BIO|HENRY|KM|KOC|LOG")) %>% #Uses text matching to keep only the columns we care about
  rename_with(~ c("Key","Name", "AOH", "BCF", "BIODEG_HL", "HenryLaw_Const", "KM", "log_KOA", "KOC", "log_KOW")) %>% #Renames the columns to be a little more readable
  rowwise() %>%
  mutate(log_KAW = (log_KOW - log_KOA), log_KOC = log10(KOC)) %>%
  select(-c(HenryLaw_Const, KOC, AOH, BCF, BIODEG_HL, KM)) %>% #Deletes unwanted columns
  ungroup() %>%
  mutate(Group = "Group2", CmpdID = row_number()) %>%
  relocate(matches("Group|CmpdID"), .before = Name)

results2 <- compounds2 %>%
  full_join(PubChem_IDs2, by = c("SMILES" = "query")) %>% #Combines previously created tables by similar columns
  full_join(InChiKeys2, by = c("cid" = "CID")) %>%
  full_join(compTox2, by = c("InChIKey" = "Key"))




resultsCombined <- results %>%
  bind_rows(results2)

print(resultsCombined)


#Optional code for creating a csv (if you want to make your graphs in Excel):

write_csv(resultsCombined, file = "Predicted_Partioning_Coefficients.csv")
```
# Part 3: Example Plots

Example plot of log KOA as a function of log KAW.

```{r plot1}
Kplot <- ggplot(data = resultsCombined, mapping = aes(x = log_KOA, y = log_KAW, color = Group, label = CmpdID))+ 
  geom_point()+ #Converts plot to a scatter plot
  geom_text()+ #Adds a text layer for labels
  labs(x = "log KOA", y = "log KAW")+ #Custom axis labels
  theme_classic() #Simplifies layout and removes gridlines
print(Kplot)
```


This is the same plot as above, but with the addition of the optional package *ggrepel*. This prevents data labels from appearing directly on top of data points.

```{r fancyplot1}
if(!require("ggrepel")) install.packages("ggrepel")
library(ggrepel)

Kplot <- ggplot(data = resultsCombined, mapping = aes(x = log_KOA, y = log_KAW, color = Group, label = CmpdID))+
  geom_point()+
  geom_text_repel(min.segment.length = 0.1)+ #Replaces 'geom_text()' in the previous code block
  labs(x = "log KOA", y = "log KAW")+
  theme_classic()
print(Kplot)

```

Another example plot, this time plotting log KAW as a function of log KOC (organic carbon:water coefficient).

```{r fancyplot2}
library(ggrepel)
Koc_plot <- ggplot(data = resultsCombined, mapping = aes(x = log_KOC, y = log_KAW, color = Group, label = CmpdID))+
  geom_point()+
  geom_text_repel(min.segment.length = 0.1)+
  labs(x = "log KOC", y = "log KAW")+
  theme_classic()
print(Koc_plot)

```
